---
title: "R for Data Science With Sports Applications"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  hiline = TRUE
)

df <- readRDS('./nba.rds')
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_duo_accent(
  primary_color = "#1381B0",
  secondary_color = "#FF961C",
  inverse_header_color = "#FFFFFF"
)
```

## Grammar of graphics: layers of a plot

- Data

- Geoms

- Aesthetic mappings

- Facets


---

### Data

- The underlying data frame.

- Think rows and columns.


---
### Geoms

- Shapes to represent the data.

- Individual geoms use one shape per row.

- Group geoms use multiple rows to create a shape.

---
### Aesthetic mappings

- Map between the aesthetic properties of a geom and column of the data.

---
### Facets

- Also called small multiples.

- Slice the rows into multiple sub populations.

---
### Identify the components of the grammar of graphics

```{r, echo=FALSE}
library(ggplot2)

# TODO: create a plot of assists vs turnovers.
plt <- ggplot(data=df, aes(x=AST, 
                    y=TOV, 
                    color=Playoff, 
                    shape=Playoff)) + 
  geom_point() + 
  labs(title="Turnovers vs. Assists",
       subtitle="2017/18 Regular Season",
       x="Assists",
       y="Turnovers")

plt
  
```

---

```{r echo=FALSE}
library(trackeR)
runZones <- zones(runs[1:4], what = "speed", breaks = c(0, 2:6, 12.5))
plot(runZones)
```

---
### Identify the components (2)

```{r echo=FALSE}
plot(runs, session = 1:3)
```


---
```{r echo=FALSE}
runSummaryFull <- summary(runs)
plot(runSummaryFull, group = c("total", "moving"),
       what = c("avgSpeed", "distance", "duration", "avgHeartRate"))
```




---

### Scatter plot: Data

- Call the `ggplot` function and pass in a data frame.
```{r}
ggplot(df)
```

---
### Scatter plot: Aesthetic mappings

- Pass the aes function with the mapping as a second argument.

```{r}
ggplot(df, aes(x=AST, y=TOV))
```
---
### Scatter plot: Geoms

- Add a layer to use one point to represent one row.


```{r}
ggplot(df, aes(x=AST, y=TOV)) + geom_point()
```

---
# More mappings:

- Shape and color
```{r}
ggplot(df, aes(x=AST, y=TOV, color=Playoff, shape=Playoff)) + 
  geom_point()

```

<!-- ### How to make a barchart -->

<!-- # ```{r} -->
<!-- # library(ggplot2) -->
<!-- #  -->
<!-- # df %>%  -->
<!-- #   arrange(-PTS) %>%  -->
<!-- #   head(5) %>%  -->
<!-- #   ggplot(aes(PTS, Team, fill=Conference)) + -->
<!-- #   geom_col()  -->
<!-- # ``` -->


---
### Lab 2

- Create a plot to understand the relationship between two pointers made and three pointers made.
- Think about the data and the aesthetic mappings that you want.
- Put it in an Rmarkdown file.